---
title: "Final Presentation"
subtitle: "Turkey Analysis: Trade Balance and Partners"
author: "Group Yellow Submarine"
institute: "MEF University"
date: "2024-01-04"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
        

---


<style>
/* Set the font family to Helvetica */
body {
  font-family: 'Helvetica', sans-serif;
}



/* Customize other elements as needed */
</style>

## Key Takeways


 **Net Importer Position:**  Import Dependency on Intermediary Goods

 **Trade Partner Concentration Risk:** 200+ Total Partners, Limited Ones on Strategic Product Categories
 
 **Product Category Risk:**  More than 80% of Total Exports/Imports coming from Top 10 Product Categories
 
 **Stagnant Revenue Streams :** Stagnant Growth in Top Export Product Categories

 **Reliance on Intermediates:** Heavy Intermediate Use, Low level of Local Input Production  

 **New Economic Model Needed:** Investments on R&D, Innovation Emphasis, Diversification of  Import & Export Sources


---
## Introduction
Group members are Sezer Türkmen, Sezgi Ayhan

Our group project, "Turkey Analysis: Trade Balance & Partners" delves into the details of Turkey's international trade landscape from 2002 to 2020.

The data, sourced from a package developed by a team led by Mauricio Vargas provides a comprehensive overview of Turkey's trade dynamics. With this study, we have;

Explored certain sectors, like textiles and machinery, significantly contribute to Turkey's export revenue. On the import side, machinery and base metals play a crucial role. 

Delved into the trends of these categories over the years, identifying fluctuations and their potential drivers.

Explored Turkey's trade partnerships, highlighting the country's diverse range of trading allies. 


```{r setup, include = FALSE}

# Necessary Libraries
knitr::opts_chunk$set(echo = TRUE)
library(htmlwidgets)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggridges)
library(knitr)
#library(kableExtra)
library(DT)

#In order to solve conflicts in different packages, we additionally loaded conflicted package to set our preference as dplyr

library(conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")

```

```{r include=FALSE}
# Load the datasets
loaded_datasets <- readRDS("wits_data.rds")
read_wits_turkey_data_only <- loaded_datasets$wits_turkey_data_only
read_wits_turkey_data_with_partners <- loaded_datasets$wits_turkey_data_with_partners
```

```{r include=FALSE, message=FALSE,warning=FALSE}
sapply(read_wits_turkey_data_only, function(x) sum(is.na(x)))
sapply(read_wits_turkey_data_with_partners, function(x) sum(is.na(x)))
```
---
## Exports & Imports Analysis



```{r fig.width=11, fig.height=5,echo=FALSE, message=FALSE,warning=FALSE}
total_exports <- read_wits_turkey_data_only %>%
  group_by(year) %>%
  summarize(total_exports = sum(trade_value_usd_exp))

total_imports <- read_wits_turkey_data_only %>%
  group_by(year) %>%
  summarize(total_imports = sum(trade_value_usd_imp))

# Visualize exports and imports over the years
ggplot() +
  geom_line(data = total_exports, aes(x = year, y = total_exports, color = "Exports"), size = 1.5) +
  geom_line(data = total_imports, aes(x = year, y = total_imports, color = "Imports"), size = 1.5) +
  labs(title = "Turkey's Exports and Imports Over the Years", x = "Year", y = "Trade Value (mn USD)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6))  # Format y-axis labels as millions




```


Turkey has experienced an upward trajectory in both exports and imports from 2002 to 2020, influenced by factors such as global crisis, commodity shocks and changes in macroeconomic policies and currency exchange rates.



---

## Net Trade Position of Turkey
```{r fig.width=11, fig.height=5, echo=FALSE,message=FALSE,warning=FALSE}

# Calculate total trade for each year
yearly_trade <- read_wits_turkey_data_with_partners %>%
  group_by(year) %>%
  summarize(total_trade = sum(trade_value_usd_exp + trade_value_usd_imp)) %>%
  arrange(desc(year)) 
  
# Calculate total export and import for each year
yearly_trade_net <- read_wits_turkey_data_with_partners %>%
  group_by(year) %>%
  summarize(total_export = sum(trade_value_usd_exp),
            total_import = sum(trade_value_usd_imp)) %>%
  arrange(desc(year))


# Calculate the net difference
yearly_trade_net <- yearly_trade_net  %>%
  mutate(trade_difference = total_export - total_import)%>%
  na.omit() 


# Visualize import and export by year
ggplot(data = yearly_trade_net, aes(x = year)) +
  geom_bar(aes(y = total_export, fill = "Export"), stat = "identity", position = "dodge", color = "darkgray", linetype = "dashed" ) +
  geom_bar(aes(y = -total_import , fill = "Import"), stat = "identity", position = "dodge", color = "darkgray", linetype = "dashed") +
  geom_bar(aes(y = trade_difference , fill = factor(sign(trade_difference))),
           stat = "identity", position = "dodge", color = "black", alpha=0.7) +
  labs(title = "Net Trade Position of Turkey",
       x = "Year", y = "Total Trade Value (mn USD)") +
  scale_fill_manual(values = c("Export" = "gray", "Import" = "lightgray",
                                "1" = "skyblue3", "-1" = "lightcoral"),
                    name = "",
                    labels = c("Net importer", "Export", "Import")) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5),
        legend.position = "top") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6))
```

Turkey has always registered net foreign trade deficit throughout the analysis period.

---


## Top Export Product Categories

```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}
export_plot_data <- read_wits_turkey_data_only %>%
  group_by(section_name) %>%
  summarize(total_export = sum(trade_value_usd_exp)/19000000000) %>%
  top_n(10)

# Truncate section names to the first 50 characters
export_plot_data$truncated_name <- str_trunc(export_plot_data$section_name, 50)

# Get distinct truncated names for colors
color_names <- unique(export_plot_data$truncated_name)

# Generate a limited number of distinct colors
colors <- brewer.pal(length(color_names), "Set3")

ggplot(data = export_plot_data, aes(x = reorder(section_name, -total_export), y = total_export, fill = truncated_name)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = paste0("$",round(total_export), "bn"),y = total_export - 3), color = "black", vjust = -0.5) +  # Add values above the columns
  labs(title = "Top 10 Diversity of Exported Products (2002-2020, Turkey)", x = "Products", y = "Annual Export Value") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 5)) +
  scale_fill_manual(values = setNames(colors, color_names), name = "")

```

Turkey's annual export revenue on average throughout the analysis period is \~125bn USD. Turkey's top 10 export revenues constitute 89% of total export income.

---
## Top Import Product Categories


```{r fig.width=11, fig.height=5,echo=FALSE, message=FALSE,warning=FALSE}
imported_plot_data <- read_wits_turkey_data_only %>%
  group_by(section_name) %>%
  summarize(total_import = sum(trade_value_usd_imp)/19000000000) %>%
  top_n(10)

# Truncate section names to the first 20 characters
imported_plot_data$truncated_name <- str_trunc(imported_plot_data$section_name, 50)

# Get distinct truncated names for colors
color_names <- unique(imported_plot_data$truncated_name)

# Generate a limited number of distinct colors
colors <- brewer.pal(length(color_names), "Set3")

ggplot(data = imported_plot_data, aes(x = reorder(section_name, -total_import), y = total_import, fill = truncated_name)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = paste0("$",round(total_import), "bn"),y = total_import- 3), color = "black", vjust = -0.5) +  # Add values above the columns 
  labs(title = "Top 10 Diversity of Imported Products (2002-2020, Turkey)", x = "Products", y = "Annual Import Value") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 5)) +
  scale_fill_manual(values = setNames(colors, color_names), name = "")

```

Turkey's import value on average throughout the analysis period is \~183 bn USD.Turkey's top 10 import value constitute 81% of total imports.

---

## Product Category Analysis

```{r fig.width=11, fig.height=5,echo=FALSE,message= FALSE}
# Calculate total trade for each section
sections <- read_wits_turkey_data_with_partners %>%
  group_by(section_name) %>%
  summarize(total_trade = sum(trade_value_usd_exp + trade_value_usd_imp)) %>%
  arrange(desc(total_trade)) %>%
  top_n(10)


# Calculate total export and import for each top partner
top_sections <- read_wits_turkey_data_with_partners %>%
  filter(section_name %in% sections$section_name) %>%
  group_by(section_name) %>%
  summarize(total_export = sum(trade_value_usd_exp),
            total_import = sum(trade_value_usd_imp)) %>%
  arrange(desc(total_export + total_import)) %>%
  top_n(10)%>%
  na.omit() 

# Calculate the net difference
top_sections <- top_sections %>%
  mutate(trade_difference = total_export - total_import)%>%
  na.omit() 

top_sections <- top_sections %>% 
  mutate(
    section_name = case_when(
      str_detect(section_name, "^Machinery and mechanical appliances") ~ "Machinery Products",
      str_detect(section_name, "^Base metals") ~ "Base Metal Products",
      str_detect(section_name, "^Natural or") ~ "Natural Products",
      str_detect(section_name, "^Vehicles,") ~ "Vehicles Equipment",
      str_detect(section_name, "^Plastics and articles") ~ "Plastics",
      str_detect(section_name, "^Product of the chemicals") ~ "Chemicals",
      str_detect(section_name,  "^Textile and textile articles") ~"Textiles",
      str_detect(section_name,  "^Vegetable Products") ~"Vegetables",
      str_detect(section_name,  "^Prepared foodstuffs") ~"Foodstuffs and Beverages",
      TRUE ~ as.character(section_name)
    )
    )
    
# Visualize import and export relations with top trading partners
ggplot(data = top_sections, aes(x = reorder(section_name, -total_export - total_import))) +
  geom_bar(aes(y = total_export , fill = "Export"), stat = "identity", position = "dodge", color = "darkgray", linetype = "dashed" ) +
  geom_bar(aes(y = -total_import , fill = "Import"), stat = "identity", position = "dodge", color = "darkgray", linetype = "dashed") +
  geom_bar(aes(y = trade_difference , fill = factor(sign(trade_difference / 1e6))),
           stat = "identity", position = "dodge", color = "black", alpha = 0.7) +
  labs(title = "Foreign Trade Surplus/Deficit by Top 10 Products",
       x = "Partner", y = "Total Trade Value (mn USD)") +
  scale_fill_manual(values = c("Export" = "gray", "Import" = "lightgray",
                                "1" = "skyblue3", "-1" = "lightcoral"),
                    name = "",
                    labels = c("Turkey net importer", "Turkey net exporter", "Export", "Import")) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5),
        legend.position = "top") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6)) 
```

---
## Trade Partnership Analysis

```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}
top_partners <- read_wits_turkey_data_with_partners %>%
  group_by(partner_name) %>%
  summarize(total_trade = sum(trade_value_usd_exp + trade_value_usd_imp)/19000000000) %>%
  arrange(desc(total_trade)) %>%
  top_n(10)

# Get a color palette from RColorBrewer
partner_colors <- brewer.pal(length(top_partners$partner_name), "Set3")

# Visualize top trading partners
ggplot(data = top_partners, aes(x = reorder(partner_name, -total_trade), y = total_trade, fill = partner_name)) +
  geom_bar(stat = "identity", color = "black") +
  geom_text(aes(label = paste0("$",round(total_trade), "bn"),y = total_trade- 3), color = "black", vjust = -0.5) +  # Add values above the columns
  labs(title = "Top 10 Trading Partners (2002-2022)", x = "Partner", y = "Annual Trade Value") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = partner_colors)+
  theme(legend.position = "none")

```

Turkey engages with more than 200 distinct export partners across its top 10 product categories, showcasing a diversified revenue stream. Top 10 trade partners' total trade value with Turkey throughout the analysis period constitute 51% of total trade value.

---
## Trade Partnership Analysis
```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}
# Calculate total trade for each partner
partners <- read_wits_turkey_data_with_partners %>%
  group_by(partner_name) %>%
  summarize(total_trade = sum(trade_value_usd_exp + trade_value_usd_imp)) %>%
  arrange(desc(total_trade)) %>%
  top_n(10)

# Calculate total export and import for each top partner
top_partners <- read_wits_turkey_data_with_partners %>%
  filter(partner_name %in% partners$partner_name) %>%
  group_by(partner_name) %>%
  summarize(total_export = sum(trade_value_usd_exp),
            total_import = sum(trade_value_usd_imp)) %>%
  arrange(desc(total_export + total_import)) %>%
  top_n(10)%>%
  na.omit() 

# Calculate the net difference
top_partners <- top_partners %>%
  mutate(trade_difference = total_export - total_import)%>%
  na.omit() 


# Visualize import and export relations with top trading partners
ggplot(data = top_partners, aes(x = reorder(partner_name, -total_export - total_import))) +
  geom_bar(aes(y = total_export, fill = "Export"), stat = "identity", position = "dodge",color = "darkgray", linetype = "dashed" ) +
  geom_bar(aes(y = -total_import, fill = "Import"), stat = "identity", position = "dodge",color = "darkgray", linetype = "dashed") +
  geom_bar(aes(y = trade_difference, fill = factor(sign(trade_difference/1e6))),
           stat = "identity", position = "dodge", color = "black", alpha=0.7) +
  labs(title = "Foreign Trade Surplus/Deficit by Top 10 Partners",
       x = "Partner", y = "Total Trade Value (mn USD)") +
  scale_fill_manual(values = c("Export" = "gray", "Import" = "lightgray",
                                "1" = "skyblue3", "-1" = "lightcoral"),
                    name = "",
                    labels = c("Turkey net importer", "Turkey net exporter", "Export", "Import")) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-6))
```

---
## Trade Partnership Analysis
```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}

# Vector of countries to filter
selected_countries <- c("Germany", "Russian Federation", "China", "Italy", "USA", "United Kingdom", "France", "Spain", "Iran", "Iraq")

top_partners <- read_wits_turkey_data_with_partners %>%
  filter(partner_name %in% selected_countries, year %in% 2015:2020) %>%
  group_by(partner_name, year) %>%
  summarize(total_export = sum(trade_value_usd_exp),
            total_import = sum(trade_value_usd_imp)) %>%
  mutate(trade_difference = total_export - total_import)


# Create a bar plot to show trade comparison with selected trading partners
ggplot(data = top_partners, aes(x = reorder(partner_name, -trade_difference), y = trade_difference/1000000, fill = factor(sign(trade_difference)))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black", height = 1.5) +  # Adjust the 'height' parameter
  facet_wrap(~year, scales = "free_y", ncol = 3) +
  scale_fill_manual(values = c("lightcoral", "skyblue3"), name = "",
                    labels = c("Turkey net importer", "Turkey net exporter")) +
  labs(title = "Trade Comparison with Top 10 Trading Partners(2015-2020)",
       x = "Partner",
       y = "Trade Difference (Export - Import, mn USD)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        legend.position = "top")
```


---
## Trade Partnership Analysis
```{r fig.width=11, fig.height=5,echo=FALSE,essage=FALSE,warning=FALSE}

# Replace "YourFirstYear" and "YourEndYear" with the actual years that is selected for the analysis
first_year <- 2015
end_year <- 2020

# Calculate the total trade value for the first and end years for each partner
total_trade_by_partner <- read_wits_turkey_data_with_partners %>%
  group_by(partner_name) %>%
  summarise(total_trade_first = sum(trade_value_usd_imp[year == first_year] + trade_value_usd_exp[year == first_year]),
            total_trade_end = sum(trade_value_usd_imp[year == end_year] + trade_value_usd_exp[year == end_year]))

# Find the top 10 partners based on the total trade value
top_10_partners <- total_trade_by_partner %>%
  top_n(10, wt = total_trade_end)

# Calculate the percentage change
percentage_change <- top_10_partners %>%
mutate(percentage_change = ((total_trade_end/total_trade_first)^(1/5)-1)*100)

# Create a lollipop chart
ggplot(percentage_change, aes(x = reorder(partner_name, percentage_change), y = percentage_change)) +
  geom_segment(aes(xend = reorder(partner_name, percentage_change), yend = 0), color = "skyblue", size = 1) +
  geom_point(color = "red", size = 3) +
  labs(title = "Top 10 Partners: 5 year CAGR growth in Total Trade Value by 2020",
       x = "Partner",
       y = "5 year CAGR growth in Total Trade Value") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

---
## Trade Partnership Analysis  by Product Categories

```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}
# Filter data for Iraq, Spain, and USA
trade_data_selected_partners<- read_wits_turkey_data_with_partners %>%
  filter(partner_name %in% c("Germany","Russian Federation","China", "United Kingdom", "Iraq", "Spain", "USA","Iran", 1))

# Add a new column for total trade
trade_data_selected_partners <- trade_data_selected_partners%>%
  mutate(total_trade_exp = (trade_value_usd_exp)/19000000)

# Group by section_name and calculate total trade values for each section_name
total_trade_values_19_year <- trade_data_selected_partners %>%
  filter(year >= 2002 & year <= 2020) %>%
  group_by(section_name, partner_name) %>%
  summarise(total_exp_value = sum(total_trade_exp))

# Filter top 5 section names
filtered_commodity_data <- total_trade_values_19_year %>%
  group_by(partner_name) %>%
  arrange(desc(total_exp_value)) %>%
  slice_head(n=5)

# Truncate section names to the first 20 characters
filtered_commodity_data$truncated_name <- str_trunc(filtered_commodity_data$section_name, 25)

# Set up the color palette
colors <- brewer.pal(length(unique(filtered_commodity_data$truncated_name)), "Set3")

# Map colors to truncated section names
color_mapping <- setNames(colors, unique(filtered_commodity_data$truncated_name))

# Use the color_mapping in ggplot
ggplot(filtered_commodity_data, aes(x = partner_name, y = total_exp_value, fill = truncated_name)) +
  geom_col() +
  scale_fill_manual(values = color_mapping) +  # Set manual color scale
  labs(title = "Annual Export Breakdown by Product Category and Partner (2002-2020)",
       x = "Partner Name", y = "Total Export mn USD",
       fill = "Sections") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))

```

---
## Trade Partnership Analysis  by Product Categories

```{r fig.width=11, fig.height=5,echo=FALSE,message=FALSE,warning=FALSE}
# Filter data for Iraq, Spain, and USA
trade_data_selected_partners<- read_wits_turkey_data_with_partners %>%
  filter(partner_name %in% c("Germany","Russian Federation","China", "United Kingdom", "Iraq", "Spain", "USA","Iran", 1))

# Add a new column for total trade
trade_data_selected_partners <- trade_data_selected_partners%>%
  mutate(total_trade_imp = (trade_value_usd_imp)/19000000)

# Group by section_name and calculate total trade values for each section_name
total_trade_values_19_year <- trade_data_selected_partners %>%
  filter(year >= 2002 & year <= 2020) %>%
  group_by(section_name, partner_name) %>%
  summarise(total_imp_value = sum(total_trade_imp))

# Filter top 5 section names
filtered_commodity_data <- total_trade_values_19_year %>%
  group_by(partner_name) %>%
  arrange(desc(total_imp_value)) %>%
  slice_head(n=5)

# Truncate section names to the first 20 characters
filtered_commodity_data$truncated_name <- str_trunc(filtered_commodity_data$section_name, 25)

# Set up the color palette
colors <- brewer.pal(length(unique(filtered_commodity_data$truncated_name)), "Set3")

# Map colors to truncated section names
color_mapping <- setNames(colors, unique(filtered_commodity_data$truncated_name))

# Use the color_mapping in ggplot
ggplot(filtered_commodity_data, aes(x = partner_name, y = total_imp_value, fill = truncated_name)) +
  geom_col() +
  scale_fill_manual(values = color_mapping) +  # Set manual color scale
  labs(title = "Average Annual Import Breakdown by Product Category and Partner (2002-2020)",
       x = "Partner Name", y = "Total Import mn USD",
       fill = "Sections") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black", size = 0.5) +  # Add horizontal line at y = 0
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 4))
```


---


## Conclusion

**Robust Growth Amid Global Shifts:**
Turkey's trade has steadily grown over two decades, navigating events like financial crises and policy changes.

**Resilience Through Deficits:**
Despite facing a persistent foreign trade deficit, Turkey has showcased its ability to navigate through economic downturns

**Key Sectors and Dependencies:**
Textiles, machinery, and base metals drive exports; machinery and base metals are crucial imports, revealing reliance on diverse intermediates. Turkey’s exports are also concentrated in the sectors where the use of imported inputs has expanded.

**Diverse Global Partnerships:**
Turkey's 200+ trade partners, including neighboring countries and major economies like Germany, Italy, China, the USA, and Russia, highlight its strategic global trade position.

**Actionable Insights:**
There is a need to enhance local input production, reduce reliance on intermediate goods, and strategically focus on R&D investments 
---

## Thank You For Listening!

[Check Out **GitHub**  for more details](https://pjournal.github.io/mef07g-yellowsubmarine)

#### Project Documents

-   [Data Preprocessing](docs/preprocesseddata.html)

-   [Exploratory Data Analysis](exploratorydataanalysis.html)

-   [Final Report](finalreport.html)

-   [Shiny App](https://sezerlucianturkmen.shinyapps.io/yellowsubmarine/)

-   [Medium Post](https://medium.com/eda-journal/a-turkey-analysis-trade-balance-partners-based-on-world-integrated-trade-systems-wits-database-ecaf9597204c)


