---
title: "Prepocessed Data"
date: "2023-11-16"
---

## Yellow Submarine's Project Proposal on WITS data

The choice of this domain stems from the significance of Turkey's role in the global trade landscape between 2002-2020. Before settling on the WITS database, we reviewed existing solutions, including the [Harvard University Atlas project](https://atlas.cid.harvard.edu/) and [Visualizations from the Observatory of Economic Complexity](https://oec.world/en).) 

We have decided on this database considering the primary sources that include:
a. Trade Partnerships: Utilize international trade databases to identify Turkey's major trade partners, assess trade volumes, and analyze the composition of exports and imports.
b. Complexity Score: Apply complexity indices to measure the sophistication of Turkey's exports and imports, shedding light on the diversification and technological intensity of its trade activities.
c. Current Account Projections: Examine historical current account data and employ forecasting models to project future trends, allowing for informed insights into Turkey's economic stability.
d. Trading Routes: Analyze shipping and logistics data to map out key trading routes, highlighting the connectivity and strategic importance of Turkey in global trade.
e. Additional Insights: Explore supplementary datasets as needed to uncover specific aspects of Turkey's trade dynamics, such as sector-specific analyses or the impact of geopolitical events.

With this project, we aim to produce a comprehensive report that synthesizes the findings into a coherent narrative. Visualizations, charts, and infographics will be employed to enhance the accessibility of the data.
---

## Dataset Installation and Preprocessing

```{r prepocesseddata,eval=FALSE, echo=TRUE,message=FALSE}

library(tidyverse)
library(dplyr)
library(ggplot2)
library(tradestatistics)
library(tibble)


wits_turkey_data_only <- ots_create_tidy_data(
  years = 2002:2020,
  reporters = "tur",
  table = "yrc"
)
head(wits_turkey_data_only)

wits_turkey_data_with_partners <- ots_create_tidy_data(
  years = 2002:2020,
  reporters = "tur",
  table = "yrpc"
)
head(wits_turkey_data_with_partners)


# Combine the datasets into a list
combined_datasets <- list(
  wits_turkey_data_only = wits_turkey_data_only,
  wits_turkey_data_with_partners = wits_turkey_data_with_partners
)

```


## Creating RDS file
```{r rdsdata,eval=FALSE, echo=TRUE}

saveRDS(combined_datasets, file = "wits_data.rds")

```
 [RDS Link](https://github.com/pjournal/mef07g-yellowsubmarine/blob/main/wits_data.rds)
 
## Monitoring the Data
```{r monitoring}
library(DT)

loaded_datasets <- readRDS("wits_data.rds")
read_wits_turkey_data_only <- loaded_datasets$wits_turkey_data_only
read_wits_turkey_data_with_partners <- loaded_datasets$wits_turkey_data_with_partners

DT::datatable(head(read_wits_turkey_data_only), editable = list(
  target = 'row', disable = list(columns = c(1, 3, 4))
))

DT::datatable(head(read_wits_turkey_data_with_partners), editable = list(
  target = 'row', disable = list(columns = c(1, 3, 4))
))


```